---
title: "Comparaison_V4_SR_LR"
author: "Paul Terzian"
date: "2026-03-27"
output: html_document
---


In this study we clustered the seq seed of V4SR filtered with the 0.005 Bokulich filter (OTU with abund > 66) with the full seq seed of extracted V4 from full18S LR.
Here we compared the abundances of OTUs in SR-only and mix clusters.


```{r setup, include=FALSE}
library(data.table)
library(ggplot2)
library(stringr)
library(stringi)
library(coin)
```

### Load tables & keep only useful column
```{r}
### nb OTU in below table is less than the total number of OTU produced by clustering because some those OTU were filtered during PANAM2's taxonomic affiliation
OTU_SR_LR <-  fread(input = "OTU_0.97_list_longformat_bokulich_0.005_extractedsacch.txt", stringsAsFactors = F, sep = "\t", col.names = "OTU", header = F)

OTU_list_metazoa <-  fread(input = "OTU_list_metazoa.txt", stringsAsFactors = F, sep = "\t", col.names = , header = F)

OTU_V4_SR <- fread(input = "OTU_distribution_tax_V4_SR_marina.txt", stringsAsFactors = F, sep = "\t")

OTU_V4_LR <- fread(input = "../Comparaison_LR_SR_V4/OTU_distribution_tax_Full18S_97_LR.txt", stringsAsFactors = F, sep = "\t")

db_taxo_SR <- OTU_V4_SR[,c(1,2,41,42,43,44)]
db_taxo_LR <- OTU_V4_LR[,c(1,2,41,42,43,44)]
db_taxo_LR$OTU_Id <- str_replace(db_taxo_LR$OTU_Id, pattern = "_LR$", "")

################## SR bokulich filter
db_abund_SR <- OTU_V4_SR[,!(c(1,2,41,42,43,44))]
db_abund_SR_id <- OTU_V4_SR[,!(c(2,41,42,43,44))]


db_abund_LR <- OTU_V4_LR[,!(c(1,2,41,42,43,44))]
db_abund_LR_id <- OTU_V4_LR[,!(c(2,41,42,43,44))]
```

### First identify which cluster are only SR or LR ## get ratio
```{r}
lrez <- lapply(OTU_SR_LR$OTU, function(x){
  
  cur_OTU <- unlist(str_split(x, pattern = ","))
  nbLR <- sum(str_count(cur_OTU, "LR"))
  nbSR <- sum(str_count(cur_OTU, "SR"))
  ratio <- nbLR/(nbLR+nbSR)*100
  cluststatus <- ifelse(test = (ratio == 100), yes = "FullLR", no = ifelse(ratio == 0, yes = "FullSR", no = "Mixed"))
  
  return(data.table(nbLR = nbLR, nbSR = nbSR, ratioLRSR = ratio, status = cluststatus))
  
})
ratiotable <- do.call(rbind, lrez)
ratiotable$OTU <- 1:nrow(ratiotable)
ratiotable[, sums := nbLR +  nbSR]

## Plot before removing singletons
hist_clusters <- ggplot(data = ratiotable, mapping = aes(x = status)) +
  geom_bar() +
  geom_text(stat='count', aes(label=..count..), vjust=0)
hist_clusters

```

## Find taxonomy for FullSR 
```{r}
# nosingletonsTable <- ratiotable[sums > 1]
# ES-ce qu'on garde que les clusters singletons ?

fullSR_otus <- OTU_SR_LR[ratiotable[status == "FullSR"]$OTU]

fullLR_otus <- OTU_SR_LR[ratiotable[status == "FullLR"]$OTU]

otu_LR_list <- str_replace(unlist(strsplit(fullLR_otus$OTU, ",")), pattern = "_LR", "")

fullmixed_otus <- OTU_SR_LR[ratiotable[status == "Mixed"]$OTU]
```


#### Clean annotation from unwanted strings
```{r}
toremove_tax <- c("environmental samples;","unclassified.+;(.+)?")

db_taxo_SR$`LCA taxonomy` <- str_remove_all(string = db_taxo_SR$`LCA taxonomy`, pattern = toremove_tax[1])
db_taxo_SR$`LCA taxonomy` <- str_remove_all(string = db_taxo_SR$`LCA taxonomy`, pattern = toremove_tax[2])

db_taxo_LR$`LCA taxonomy` <- str_remove_all(string = db_taxo_LR$`LCA taxonomy`, pattern = toremove_tax[1])
db_taxo_LR$`LCA taxonomy` <- str_remove_all(string = db_taxo_LR$`LCA taxonomy`, pattern = toremove_tax[2])
```


## Find unique longest taxonomic affiliations in full SR clusters
```{r}
otu_SR_list <- str_replace(unlist(strsplit(fullSR_otus$OTU, split = ",")), pattern = "_SR", "")

only_SR_taxo <- db_taxo_SR[OTU_Id %in% otu_SR_list]
uniq_timmed_taxo <- str_replace(unique(only_SR_taxo$`LCA taxonomy`), " ;", "")

unique_taxa_SR <- data.table(taxoref = uniq_timmed_taxo)
```


## Find unique longest taxonomic affiliations in Mixed clusters
```{r}
otu_mixed_list <- unlist(strsplit(fullmixed_otus$OTU, split = ","))

otu_mixed_list_SR <- str_replace(otu_mixed_list[grepl("SR", otu_mixed_list)], pattern =  "_SR", "")
otu_mixed_list_LR <- str_replace(otu_mixed_list[grepl("LR", otu_mixed_list)], pattern = "_LR", "")

only_mixed_taxo_SR <- db_taxo_SR[OTU_Id %in% otu_mixed_list_SR]
only_mixed_taxo_LR <- db_taxo_LR[OTU_Id %in% otu_mixed_list_LR]

only_abund_taxo_SR <- db_abund_SR_id[OTU_Id %in% otu_mixed_list_SR]
only_abund_taxo_SR_fullSR <- db_abund_SR_id[OTU_Id %in% otu_SR_list]


only_abund_taxo_LR_mixed <- db_abund_LR_id[OTU_Id %in% otu_mixed_list_LR]
only_abund_taxo_LR_fullLR <- db_abund_LR_id[OTU_Id %in% otu_LR_list]

missingfullSRinOTUtable <- otu_SR_list[which(!otu_SR_list %in% db_abund_SR_id$OTU_Id)]

only_mixed_taxo <- rbind(only_mixed_taxo_SR, only_mixed_taxo_LR)

uniq_timmed_taxo <- str_replace(unique(only_mixed_taxo$`LCA taxonomy`), " ;", "")

unique_taxa_mixed <- data.table(taxoref = uniq_timmed_taxo)
```


## Plot the whole distribution
```{r}
sumrow_mixed <- rowSums(only_abund_taxo_SR[,!1])
sumrow_fullSR <- rowSums(only_abund_taxo_SR_fullSR[,!1])

thetable <- data.table(c(only_abund_taxo_SR$OTU_Id, only_abund_taxo_SR_fullSR$OTU_Id), c(sumrow_mixed, sumrow_fullSR), c(rep("mixed", nrow(only_abund_taxo_SR)), rep("Full_SR", nrow(only_abund_taxo_SR_fullSR))))
colnames(thetable) <- c("OTU_id", "abundance", "type")

thetable_ordered <- thetable[order(-abundance)]

test <- wilcox.test(abundance ~ type, data = thetable_ordered)

hist_colored_log10 <- ggplot(thetable_ordered, aes(x = reorder(OTU_id, -abundance),y=abundance, fill = type)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values=c("#914e50","#b4c1c8")) +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank()) +
  scale_y_log10() + ggtitle(paste("Clusters of OTU sorted by abundance", test$p.value))

thetable_ordered[type == "mixed", mean(abundance)]
thetable_ordered[type == "Full_SR", mean(abundance)]
```

