---
title: "Find V4 from aln"
author: "Paul Terzian"
date: "2026-03-27"
output: html_document
---


* This script was design to perform the V4 extraction from 18S sequences, as part of the analysis described in *Profiling the eukaryotic diversity over a year in a lake ecosystem through short- and long-read metabarcoding*.

* It take as input the output of the mothur clustering command line and output the V4 regions.

### Loading data and libraries 
```{r setup}
library(stringr)
library(Biostrings)
library(data.table)

maln <- readDNAStringSet("output_alignment.fasta")
```


### Run through aligned sequences
```{r}
iter <- seq(from=1, to = 4250, by = 2)

lV4 <- lapply(iter, function(x){
  
  cur_ref <- unlist(strsplit(as.character(maln[[x]]), ""))
  nogaps_position <- which(cur_ref != "-")
  start_pos <- nogaps_position[579]
  end_pos <- nogaps_position[936]
  query <- maln[[x+1]]
  query_trimmed <- DNAStringSet(query, start = start_pos, end = end_pos)
  return(query_trimmed)
  
})

names(lV4) <- names(maln[iter+1])
names(lV4)
# 
largeDNAstringSet <- DNAStringSet(sapply(lV4, `[[`, 1))

mean(sapply(largeDNAstringSet, FUN=length))

writeXStringSet(largeDNAstringSet, filepath = "V4_extracted_fromSaccharomyces.fasta",format = "fasta")
```



